Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-03 how.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 \usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item Software defects in certain object-oriented JAVA systems listed in \tion{oo}:  Ant Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces;
\item The runtime of the software systems discussed in \tion{cpu}: Berkeley DB, Apache, SQLite, LLVM, 
  x264.
\ei
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. This is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}. 
We find   HOW's instance-based approach useful since it can be applied
to domains that lack a model, but possess a historical log of observed behaviours.

@@

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of software defects and software runtimes (at least, for the data sets studied here).
\ee
The rest of this paper is structured as follows.

\section{From ``Prediction'' to ``Planning''}
A major  problem with building a planning mechanism in software engineering is ``how to verify the generated
plans?''. For example,  in the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

@@

We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains with data, from which data miners can build
predictors of adequate performance, then those predictors can assess the value
of the changes proposed by HOW.
As shown in \fig{work}, 
project data is divided into two disjoint sets {\em train} and {\em test}
(so {\em train} $\cup$ {\em test} $=\;\emptyset$).
Next, from the train set, we build both a {\em planner} as well
as a {\em predictor}. Our general framework does not   commit to any particular  choice
of {\em planner} or {\em predictor}. For the purposes of this discussion 
our {\em planner}
will be HOW (discussed below)and our
our {\em predictors} will be the Random Forest and Random Forest
Regressor\footnote{Decision tree learners find the attribute whose value's
select for for sections of the data with least variation in the class
variable. Decision tree learners then recurse on each selection~\cite{breiman84}. A random forest is a set of decision trees, each built
from some random subset of the rows and columns. The conclusions
of the forest are a ``vote'' generated by all the trees~\cite{Breiman2001}. 
A standard Random Forest predicts for a discrete class variable while
a Random Forest Regressor generates numeric predictions.} taken from the SciKit
Learn toolkit~\cite{Pedregosa2012}.

Turning now to the {\em test} data, this is based to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta.

\subsection{When Not to Plan: Problem of Poor Planner}

The clear advantage of this approach is that it can be readily and automatically
applied to domains with historical data. That said, the drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor planners}, it is wise to demand high standards form the {\em
satisfactory} threshold used in this \fig{work}.

\begin{figure}[!t]
\small
\[  
\begin{array}{r}
project\\
data
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{e.g.\;Random\;Forest}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{e.g.\; HOW}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{after} /  \mathit{before}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

How prevalent is the problem of poor planners? While
this is hard to answer in the general case, we can say that for the 17 data
sets explored in this study, only 3 suffer from this problem.
This study's    data   from two  types of domains:
\bi
\item
In one type, we 
build plans to change static code attributes in order to reduce defects in
Java classes.  Here,,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything over   70\%. 
\item 
In the other type, we   build
plans to change configurations settings within a Makefile such that the resulting compiled
code will run faster. Here, performance can be measured in terms of difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 1- \frac{abs(a - p)}{a}$.
Using engineering judgement, we define satisfactory $s$ as anything over   70\%. 
\ei
This paper explores 12 static code data sets
Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.} and five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. For each of these data sets,
  we generated 25 {\em before} scores from
  random split train:test splits in the ratio 4:1 train.
 The median {\em before} scores from  the static code data sets were:
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
The median {\em before}   scores from  the configuration data were:
%issue 24
\[s=\{98,97,95,88,92\}\]
Note that we need to abort planning in only $3/17$ of our data sets.

\subsection{Are These Plans ``Causal''?}
With one small variation,  the above framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality~\cite{Hill1965}
and Granger's G-casualty~\cite{granger80}. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including:
\be
\item
Plausibility;
\item
Strength of the effect; 
\item
Specificity: locality of the effect
to specialized groups;
\item
Support by experimental evidence;
\item
Coherence between cause-effects noted
in the field and in laboratory samples;
\item
Temporally: the effect has to follow the cause;
\item
Analogy: similar causes have similar effects;
\item
Biological gradient: more exposure to the cause leads to more of an
effect)
\item
Consistency: the same cause-effect pairings have been seen by different persons
in different places.
\ee
It is hard to demonstrate H-causality in software engineering.
For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is far easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects).

The practical difficulties associated with Hill's criteria lead Granger~\cite{granger80}
to propose another criteria\footnote{Granger  notes that
his proposal was based on earlier work by Norbert Weiner~\cite{Seth2007}.}.
G-causality is based on building predictions based on {\em past} data,
then showing that this works for {\em future} data~\cite{granger80}.

Note that the standard 10-way cross-validation
procedure  in data mining does {\em not} demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
Cross-validation is then repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

Jiang et al.~\cite{me11f} and Lumpe et al.~\cite{me11f} propose a modification to cross-validation which lets them
demonstate G-causal defect effects in softwrare engineering.
They divide project data  into $N$ bins, each of with contain data collected at similar times.
They then sort those bins (on the collection date) and use bins $i,j$ to predict for $k$ where 
$i<j<k$.

We adapt the work of Jiang,Lumpe et al.  as follows. In  \fig{work}, we demand that the {\em test}
data is not collected before the {\em train} data.  With that change,
we can reasonably claim that the plan's generating by HOW are G-causal.

>>>>

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 \usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}

\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item The software defects found in  twelve object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the six software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
Many papers use this above data to build {\em predictors} for defects and runtimes.
However, none of those papers then go on to use this data to build {\em planners}.

HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. This is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}. 
We find   HOW's instance-based approach useful since it can be applied
to domains that lack a model, but possess a historical log of observed behaviours.

@@

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\ee
The rest of this paper is structured as follows.

\section{From ``Prediction'' to ``Planning''}
A major  problem with building a planning mechanism in software engineering is ``how to verify the generated
plans?''. For example,  in the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be mnaually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

@@

We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains with data, from which data miners can build
predictors of adequate performance, then those predictors can assess the value
of the changes proposed by HOW.
As shown in \fig{work}, 
project data is divided into two disjoint sets {\em train} and {\em test}
(so \mbox{{\em train} $\cup$ {\em test} $=\;\emptyset$}).
Next, from the train set, we build both a {\em planner} as well
as a {\em predictor}. Our general framework does not   commit to any particular  choice
of {\em planner} or {\em predictor}. For the purposes of this discussion 
our {\em planner}
will be HOW (discussed below)and our
our {\em predictors} will be the Random Forest and Random Forest
Regressor\footnote{Decision tree learners find the attribute whose value's
select for for sections of the data with least variation in the class
variable. Decision tree learners then recurse on each selection~\cite{breiman84}. A random forest is a set of decision trees, each built
from some random subset of the rows and columns. The conclusions
of the forest are a ``vote'' generated by all the trees~\cite{Breiman2001}. 
A standard Random Forest predicts for a discrete class variable while
a Random Forest Regressor generates numeric predictions.} taken from the SciKit
Learn toolkit~\cite{Pedregosa2012}.

Turning now to the {\em test} data, this is passed to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta.

\begin{figure}[!t]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{after} /  \mathit{before}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

\subsection{Are These Plans ``Causal''?}
With one additional stipulation,   this framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality~\cite{Hill1965}
and Granger's G-casualty~\cite{granger80}. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including
{\em  plausibility};
{\em strength} of the effect;
{\em 
specificity}, i.e. locality of the effect
to specialized groups;
{\em
support} by experimental evidence;
{\em 
coherence} between cause-effects noted
in the field and in laboratory samples;
{\em
temporally} i.e. the effect has to follow the cause;
{\em 
analogy}, i.e.  similar causes have similar effects;
{\em
biological gradient}, i.e.  more exposure to the cause leads to more of an
effect; and
{\em
consistency}, i.e.  the same cause-effect pairings have been seen by different persons
in different places.

It is hard to demonstrate H-causality in software engineering.
For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is far easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects). In the latter,
effects are more {\em probabilistic} than {\em categorical}-- by which we mean that 
some cause tends to cause some effect at some probability $p$ less that one.
This in turn means that at some probability $q$ (where $0 \le q < p \le 1$)
that we will find examples of {\em inconsistency} where some cause does {\em not} lead
to some effect (thus violating one of Hill's criteria for demonstrating causality).

The  difficulties associated with Hill's criteria lead Granger~\cite{granger80}
to propose another more practical criteria\footnote{Granger  himself comments that
his proposal was based on earlier work by Norbert Weiner~\cite{Seth2007}.}.
\mbox{G-causality} is based on building predictions based on {\em past} data,
then showing that this works for {\em future} data~\cite{granger80}. 
In should be noted  that the standard 10-way cross-validation
procedure used  in data mining does not necessarily demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
This is repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

Jiang et al.~\cite{me11f} and Lumpe et al.~\cite{me11f} propose a modification to cross-validation.
which lets a researcher
demonstrate G-causal defect effects in software engineering.
In this modification, we  divide project data  into $N$ bins, each of with contain data collected at similar times.
They then sort those bins (on the collection date) and use bins $i,j$ to build models that
are applied to bin  $k$ where 
$i<j<k$. Note that this procedure ensures that we are building theories on data seen
before the test data.

We adapt the work of Jiang,Lumpe et al.  as follows. In  \fig{work}, we demand that the {\em test}
data is not necessarily collected before the {\em train} data:
\bi
\item[\#1:] For data collected from software released in multiple versions, we should hence use  versions  $i,j$ to build models that are applied to version  $k$ where $i<j<k$.
\item[\#2:] For data collected from Monte Carlo simulations (where inputs are selected at random)
then as long as simulation $i$ is not used to adjust simulation $j> i$, then any item in that
data could be generated before any other. For this kind of data, we  can use the first $X\%$
of the simulation for training, and the remaining data for testing.
\ee
In the two case studies of this paper, the Jureczko object-oriented defect data comes from software
released in multiple versions while the Seigmund configuration data comes from Monte Carlo simulations.
Hence, we use procedure \#1 for the defect data and procedure \#2 for the configuration data.

With these changes,
we can reasonably claim that the plan's generating by HOW are G-causal.

\subsection{When Not to Plan: the Problem of Poor Predictors}

The clear advantage of the approach of \fig{work} is that it can be readily and automatically
applied to domains with historical data. That said, the obvious drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor predictors}, it is hence wise to demand high standards for the {\em
satisfactory} threshold used in this \fig{work}.

How prevalent is the problem of poor planners? 
In the general case, this is hard to answer. However, as discussed in this section, for the 17 data
sets explored in this study, only 
%XXX 
3 suffer from this problem.

This study's data from two types of domains: the Jureczko   object-oriented data  
and the Seigmund configuration data. 
For the Seigmund data, the plans generated by HOW advise on how  to change the   configurations settings within a Makefile
(in order to make resulting compiled
code  run faster). Here, performance can be measured in terms of  difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 100*(1- \frac{abs(a - p)}{a})$.
Using engineering judgement, we define satisfactory $s$ as anything over   70\%. 
This paper  explores    six Seigmund configuration data sets:  Berkeley DB (Java and Windows versions), Apache, SQLite, LLVM, and
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. 
  After splitting that data into equal sized train:test groups, a Random Forest
  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of
%issue 24
\[s=\{99.9, 99.8, 99.4, 99.1, 96,1\}\]
(Aside: if the reader doubts that such high scores are achievable, we note that these scores are 
consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)

For the  Jureczko   object-oriented data,
the plans generated by HOW advise   how to change static code attributes in order to reduce defects in
Java classes.  Here,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything over   70\%. 
This paper explores 12 of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.}.
For each of these data sets, we generated  {\em before} scores  by finding the latest
version $k$, then training  a predictor on versions $k-2$ and $k-1$. As reported below,
our initial recall scores from this data were unpromising (very low recall values). However,
after applying certain data pre-processing steps (described below), a Random Forest generated
on the $k-2,k-1$ versions achieved recalls on version $k$ of
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
From the above, we conclude that all the Seigmund configuration data and most of the Jureczko data
avoids the problem of poor predictors. Hence, we can use that data for the planning
experiments described below.

\section{Experimental Methods}

blah blah data. algorithms. statistics
